Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
edit

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We explore a range of Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations that manages to reproduce the line broadening
and flux change at the line's center within a $7\%$ and $30\%$
precision, respectively.
\end{abstract}

@@

\begin{figure*}
\begin{center}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
\end{center}
\caption{\textbf{Standard Deviation trends.} Results for all the
  Radiative Transfer simulations (in triangles) compares against the
  Doppler Shift model (lines).
  \label{fig:standard_deviation}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
\end{center}
\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
    Transfer and the lines represent the Doppler Shift.
    \label{fig:skewness}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
\end{center}
\caption{\textbf{Bimodality for RT and DS:} The triangles represent the Radiative
	Transfer and the lines represent the Doppler Shift.
	\label{fig:bimodality}}
\end{figure*}

@@

\subsubsection{Standard Deviation}
Figure \ref{fig:standard_deviation} summarizes the standard deviation
results for all our models.
Each panel shows the STD as a function of \vrot.
All panels were computed using a viewing angle of $\theta =
90^{\circ}$ (perpendicular to the rotation axis), which has the most
extreme ifluence from rotation.
 The black triangles
correspond to the full RT solution and the line to the DS
approximation.  
The optical depth increases from top to bottom and the outflow
velocity from left to right.
This quantitative plot confirms that the line width increases with
rotational velocity and  optical depth.
The DS successfuly reproduces all trends with the optical depth,
rotational velocity and outflow velocity.
However, the DS consistently underestimates the STD. 
The difference between the RT and DS increases with the outflow
velocity and the rotational velocity, and decreases with increasing
optical depth.
In the range of parameter space explore, this difference has as an
upper bound of $\sim 7\%$, $3\%$ and $\sim 2\%$ for
 $\tauh=10^5$, $10^6$ and $10^7$, respectively.

@@

Figure \ref{fig:skewness} presents the skwewness results for all the
models together with the DS comparison following the same layout as
Figure \ref{fig:standard_deviation}.
In all cases the skewness is negative showing that all the lines
are unbalanced towards the red side of the spectrum.
Skewness increases with rotational velocity and decreases with
optical depth; rotation tries to smoothes the line diminishing the
asymetries while a higher optical depth reinforces the line asymetries.
The skewness does not have a monotonous trend with outflow velocity because
there is a transition between double and single peak line; for low
outflow velocities the skewness signals the balance between the two
existing peaks while for high outflow velocities it quantifies the
asymmetry of the already dominant read peak.
The DS reproduces the main trends, again with an underestimation that
decreases at higher optical depths and increases with larger values of
the rotational velocity and outflow velocity.
In this case the differences between RT and DS have an upper bound of
$85\%$, $35\%$ and $5\%$ 
for  $\tauh=10^5$, $10^6$ and $10^7$,
respectively.

@@

Figure \ref{fig:bimodality} shows the results for the bimodality using
the same layout as in the two previous Figures.
Following the reasoning about the skewness, we observe that
increasing the outflow velocity increases the value of bimodality,
that is, it transitions to a more pronounced single peak. 
The trend as a function of the rotational velocity and the optical
depth are not monotonous.
When the outflow velocity is low (\vout$<50\ \kms$), an increasing
rotational velocity smears the two asymmetrical peaks pushing the line
morphology towards a single peaks, making the bimodality statistics
increase. 
On other situations (\vout$=50\ \kms$ and \tauh$\geq 10^6$) higher
rotational velocities the bimodality statistics decreases, which means
that it manages to slightly enlarge the already dominant red peak.
The DS continous to reproduce the main trends, still underestimating
the bimodality statistics.
As expected from the previous results the difference betwen RT and DS
decreases at higher optical depths and increases with increasing
values of the rotational and outflow velocities.
In this case the differences have an upper bound of $4\%$, $2\%$ and
$1\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.

\subsection{Intensity at line's center}

In Figure \ref{fig:valley_intensity} we quantify how the intensity at
the line's center (i.e. the valley) changes with the viewing angle,
the outflow velocity and the optical depth.
These results correspond to a fixed rotational velocity of
\vrot$=100\ \kms$.
The triangles correspond to the RT simulations and the line represents
the DS results. 
The valley intensity is expresed as a fraction of the maximum peak
intensty in the line, as such the valley/peak ratio is always less
than one. 
In every panel we see that the valley/peak ratio decreases as the
observer moves from a line  of sight perpendicular to the rotation
axis onto a parallel line of sight. 
This is a clear demonstration of the viewing angle dependency
introduced by rotation.
The valley/peak ratio at $\cos\theta=1$ matches results
without rotation, this shows that for increasing rotational velocity
the valley/peak ratio increases.
In turn, for increasing optical depth or outflowing velocity this
ratio decreases.
Again, the DS results correctly follow the trends for the full RT
simulations.

@@

\subsection{Theoretical Insights}
\color{red}
Why Doppler Shift works. Talk about radiative transfer...
\color{black}
In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different
sides of the galaxy. As the LAE is rotating, one side is being redshifted while
the other is blueshifted. We see that the combined spectrum is a weighted line,
in solid black, that combines these past two. This shows that there is in fact a
Doppler Shift and then it justifies our choice to apply it after the outflows spectrum is
obtained.

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We explore a range of Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations that manages to reproduce the line widths
and flux change at the line's center within a $7\%$ and $50\%$
precision, respectively.
\end{abstract}

@@

\begin{figure*}
\begin{center}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
\end{center}
\caption{\textbf{Standard Deviation trends.} Results for all the
  Radiative Transfer simulations (in triangles) compares against the
  Doppler Shift model (lines).
  All panels correspond to a viewing angle of $\theta = 90^{\circ}$
  (perpendicular to the rotation axis). 
  The optical depth increases from top to bottom and the outflow
  velocity from left to right.
  \label{fig:standard_deviation}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
\end{center}
\caption{\textbf{Skewness trends.} Results for all the
  Radiative Transfer simulations (in triangles) compares against the
  Doppler Shift model (lines).
  Follows the same layout as Figure \ref{fig:standard_deviation}. 
  \label{fig:skewness}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
\end{center}
\caption{\textbf{Bimodality trends.} Results for all the
  Radiative Transfer simulations (in triangles) compares against the
  Doppler Shift model (lines). 
  Follows the same layout as Figure \ref{fig:standard_deviation}. 
  \label{fig:bimodality}}
\end{figure*}

@@

\subsubsection{Standard Deviation}
Figure \ref{fig:standard_deviation} summarizes the standard deviation
results for all our models.
Each panel shows the STD as a function of \vrot.
All panels were computed using a viewing angle of $\theta =
90^{\circ}$ (perpendicular to the rotation axis), which has the most
extreme ifluence from rotation.
 The black triangles
correspond to the full RT solution and the line to the DS
approximation.  
The optical depth increases from top to bottom and the outflow
velocity from left to right.
This quantitative plot confirms that the line width increases with
rotational velocity and optical depth.
These trends are expected; higher rotational velocities can be seen as
an addition of different Doppler shifts that smear out the line, while
a higher optical depth translates into a larger number of scatterings
that increase the probability of a photon to diffuse in frequency
resulting in a broader line.

The DS successfuly reproduces all trends with the optical depth,
rotational velocity and outflow velocity.
However, the DS consistently underestimates the STD. 
The difference between the RT and DS increases with the outflow
velocity and the rotational velocity, and decreases with increasing
optical depth.
In the range of parameter space explore, this difference has as an
upper bound of $\sim 7\%$, $3\%$ and $\sim 2\%$ for
 $\tauh=10^5$, $10^6$ and $10^7$, respectively.

@@

Figure \ref{fig:skewness} presents the skwewness results for all the
models together with the DS comparison following the same layout as
Figure \ref{fig:standard_deviation}.
In all cases the skewness is negative showing that all the lines
are unbalanced towards the red side of the spectrum.
Skewness increases with rotational velocity and decreases with
optical depth; rotation tries to smoothes the line diminishing the
asymetries while a higher optical depth reinforces the line asymetries.
The skewness does not have a monotonous trend with outflow velocity because
there is a transition between double and single peak line; for low
outflow velocities the skewness signals the balance between the two
existing peaks while for high outflow velocities it quantifies the
asymmetry of the already dominant read peak.

The DS reproduces the main trends, again with an underestimation that
decreases at higher optical depths and increases with larger values of
the rotational velocity and outflow velocity.
In this case the differences between RT and DS have an upper bound of
$85\%$, $35\%$ and $5\%$ 
for  $\tauh=10^5$, $10^6$ and $10^7$,
respectively.

@@

Figure \ref{fig:bimodality} shows the results for the bimodality using
the same layout as in the two previous Figures.
Following the reasoning about the skewness, we observe that
increasing the outflow velocity increases the value of bimodality,
that is, it transitions to a more pronounced single peak. 
The trend as a function of the rotational velocity and the optical
depth are not monotonous.
When the outflow velocity is low (\vout$<50\ \kms$), an increasing
rotational velocity smears the two asymmetrical peaks pushing the line
morphology towards a single peaks, making the bimodality statistics
increase. 
On other situations (\vout$=50\ \kms$ and \tauh$\geq 10^6$) higher
rotational velocities the bimodality statistics decreases, which means
that it manages to slightly enlarge the already dominant red peak.

The DS reproduces the main trends while underestimating the bimodality
statistics. 
As expected from the previous results the difference betwen RT and DS
decreases at higher optical depths and increases with increasing
values of the rotational and outflow velocities.
In this case the differences have an upper bound of $4\%$, $2\%$ and
$1\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.

\subsubsection{Intensity at line's center}

In Figure \ref{fig:valley_intensity} we quantify how the intensity at
the line's center (i.e. the valley) changes with the viewing angle,
the outflow velocity and the optical depth.
These results correspond to a fixed rotational velocity of
\vrot$=100\ \kms$.
The triangles correspond to the RT simulations and the line represents
the DS results. 
The valley intensity is expresed as a fraction of the maximum peak
intensty in the line, as such the valley/peak ratio is always less
than one. 
In every panel we see that the valley/peak ratio decreases as the
observer moves from a line  of sight perpendicular to the rotation
axis onto a parallel line of sight. 
This is a clear demonstration of the viewing angle dependency
introduced by rotation.

The valley/peak ratio at $\cos\theta=1$ matches results
without rotation, this shows that for increasing rotational velocity
the valley/peak ratio increases.
In turn, for increasing optical depth or outflowing velocity this
ratio decreases.
Once again, the DS results correctly follow the trends for the full RT
simulations.  
This time the differences have an upper bound of $55\%$, $2\%$ and
$1\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.

@@

\subsection{Theoretical Insights}

In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different
sides of the galaxy. As the LAE is rotating, one side is being redshifted while
the other is blueshifted. We see that the combined spectrum is a weighted line,
in solid black, that combines these past two. This shows that there is in fact a
Doppler Shift and then it justifies our choice to apply it after the outflows spectrum is
obtained.
